Reanalysis Step 5: Main Effects Outcome Analysis

### 2025-08-31 reanalysis-5-outcome_analysis.qmd
fix

    # Calculate AME with clustering and weights
    # marginaleffects handles survey designs well
    # the model already handled the clustering
    ame <- marginaleffects::avg_comparisons(
      model,
      variables = "ever_lapse_binary",
      type = "response"   # average difference in predicted probabilities
    )

@@

  # Calculate and display the contrast (risk difference)
  contrasts_lapse <- pairs(emm_results)
  contrasts_df <- as.data.frame(contrasts_lapse)

  kable(contrasts_df |> 
        mutate(
          Contrast = "Lapsed - Not Lapsed",
          Risk_Difference = round(estimate, 3),
          SE = round(SE, 4),
          CI_Lower = round(lower.CL, 3),
          CI_Upper = round(upper.CL, 3),
          p.value = format.pval(p.value, digits = 4, eps = 0.0001)
        ) |>
        select(Contrast, Risk_Difference, SE, CI_Lower, CI_Upper, p.value),
        caption = "Main effect: Risk difference with proper variance estimation",
        col.names = c("Contrast", "Risk Difference", "SE", "Lower CI", "Upper CI", "P-value"))
} else {
  # If emmeans failed, display message
  print("Note: Predicted probabilities could not be calculated using emmeans.")
  print("The main analysis results (pooled coefficients and AMEs) above remain valid.")
}
```

>>>>

    # Calculate AME with clustering and weights
    # marginaleffects handles survey designs well
    # the model already handled the clustering
    ame <- marginaleffects::avg_comparisons(
      model,
      variables = "ever_lapse_binary",
      type = "response",   # average difference in predicted probabilities
      newdata = imp_data,
      wts = "weights"
    )

@@

  # Calculate and display the contrast
  contrasts_lapse <- pairs(emm_results)
  contrasts_df <- as.data.frame(contrasts_lapse)

  # When emmeans is on response scale, pairs() gives odds ratios
  # Calculate risk difference from the predicted probabilities
  risk_diff <- pred_df$Predicted_Prob[pred_df$Lapse_Status == "Lapsed"][1] - 
               pred_df$Predicted_Prob[pred_df$Lapse_Status == "Not Lapsed"][1]

  # Display odds ratio from contrast
  kable(contrasts_df |> 
        mutate(
          Contrast = "Lapsed vs Not Lapsed",
          Odds_Ratio = round(odds.ratio, 3),
          SE_OR = round(SE, 4),
          CI_Lower = round(asymp.LCL, 3),
          CI_Upper = round(asymp.UCL, 3),
          p.value = format.pval(p.value, digits = 4, eps = 0.0001)
        ) |>
        select(Contrast, Odds_Ratio, SE_OR, CI_Lower, CI_Upper, p.value),
        caption = "Main effect: Odds ratio for lapse effect",
        col.names = c("Contrast", "Odds Ratio", "SE", "Lower CI", "Upper CI", "P-value"))

  # Also display risk difference
  risk_diff_df <- data.frame(
    Contrast = "Lapsed - Not Lapsed",
    Risk_Difference = round(risk_diff, 3),
    stringsAsFactors = FALSE
  )

  kable(risk_diff_df,
        caption = "Main effect: Risk difference (from predicted probabilities)",
        col.names = c("Contrast", "Risk Difference"))
} else {
  # If emmeans failed, display message
  print("Note: Predicted probabilities could not be calculated using emmeans.")
  print("The main analysis results (pooled coefficients and AMEs) above remain valid.")
}
```
